---
title: "Analysis of fit"
author: "Blain Morin"
date: "December 18, 2017"
output: pdf_document
---

$$ Y_{i} = \alpha _{j[i]}  + \sum \beta _{p}X_{pi} + \epsilon_{i} , for hospitals (i=1 to N)$$
$$ \alpha _{j} = a + \sum b_{k}W_{kj} + u_{j}, for markets (j = 1 to J)$$


```{r, results='hide'}
### load required packages
library(lme4)
library(tidyverse)
### rstan requires having rtools installed
library(rstan)

###read in data
k12ReducedRG = read_csv("k12ReducedRG.csv")
```


First, we create the null model with no fixed effects and a market random effect. This preliminary step allows us to partition the variance into hospital and market levels. 

```{r}

### This is the code for a null model with one random effect (hospital referral region) 
m1 = lmer(episode ~ 1 + (1 | hrr) , data = k12ReducedRG)

summary(m1)

```

To check if this null model appoximates our data, we can use STAN to draw samples. We use the parameters found above to specify our STAN code.

```{r, cache=TRUE, results='hide'}
###prepare data for STAN

m1_stan_data = list(episode = k12ReducedRG$episode,
               hrr = as.integer(as.factor(k12ReducedRG$hrr)),
               K = length(unique(k12ReducedRG$hrr)))

### Specify prior parameters
N = nrow(k12ReducedRG)

### Draw samples with stan
stan_null = stan(file = "m1.stan", data = m1_stan_data, iter = 2000, chains = 4)


```

```{r}

print(stan_null, pars = c("beta", "sigma_e", "sigma_w"),
      probs=c(0.025, 0.5, 0.975))



```

The mean intercept in the stan model (18987) is similar to the observed data model (18912).

---

Next, we add the hospital level predictors.

```{r}
###rename columns 
k12ReducedRG = k12ReducedRG %>%
  rename(cmirank = "Rank for Variable cmi",
         dshpctrank = "Rank for Variable dshpct",
         mdadjadmitrank = "Rank for Variable mdadjadmit")

m2 = lmer(episode ~ qstarrating +
        qieffort +
        jchaoaccredited +
        urbanlocation +
        mdaffiliation +
        ownershipstatus +
        hospitalbedsize +
        cmirank +
        dshpctrank +
        mdadjadmitrank +
        (1 | hrr),
        data = k12ReducedRG)

summary(m2)

```


Next, we check the model using STAN. 

```{r, results='hide'}
###change data to only complete cases
k12ReducedRG = k12ReducedRG[complete.cases(k12ReducedRG),]

###change N
N = nrow(k12ReducedRG)

### prepare data for STAN
m2_stan_data = list(episode = k12ReducedRG$episode,
               hrr = as.integer(as.factor(k12ReducedRG$hrr)),
               qstar = k12ReducedRG$qstarrating,
               qieffort = k12ReducedRG$qieffort,
               accredited = k12ReducedRG$jchaoaccredited,
               urban = k12ReducedRG$urbanlocation,
               mdaffiliation = k12ReducedRG$mdaffiliation,
               ownership = k12ReducedRG$ownershipstatus,
               bedsize = k12ReducedRG$hospitalbedsize,
               cmi = k12ReducedRG$cmirank,
               dsh = k12ReducedRG$dshpctrank,
               mdadjadmitrank = k12ReducedRG$mdadjadmitrank,
               K = length(unique(k12ReducedRG$hrr)),
               id = as.integer(as.factor(k12ReducedRG$Provider)))



### run STAN simulation
m2_stan = stan(file = "m2.stan", data = m2_stan_data, iter = 1000, chains = 2)

```

```{r}
print(m2_stan, pars = c("beta", "sigma_e", "sigma_w"),
      probs=c(0.025, 0.5, 0.975))

summary(m2)
```